---
title: "Policy Effects for Income less than $35,000"
---

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(acs)
library(tidycensus)
library(tigris)
library(plotly)
library(shiny)
library(ggplot2)
library(viridis)
library(gt)
weights_geo_inc <- read.csv("img/weights_geo_inc.csv")
geo_census <- get_acs(geography = "tract",
                                  year = 2020,
                                  variables = c(Total = "B22003_001"
                                  ),
                                  state = 51, # VA
                                  county = 059,
                                  survey = "acs5",
                                  geometry = TRUE)
geo_census$GEOID <- as.numeric(geo_census$GEOID)
fairfax <- geo_census[, c(1, 2, 6)]
weights_geo_inc <- weights_geo_inc %>% left_join(fairfax, by = c("geoid"="GEOID"))
#View(weights_geo_inc)

names(weights_geo_inc)[names(weights_geo_inc)=='monthly_mean_COVID_benefit'] <- 'Monthly mean COVID benefit'
names(weights_geo_inc)[names(weights_geo_inc)=='monthly_mean_SNAP_benefit'] <- 'Monthly mean SNAP benefit'
```

## Distributions of COVID and SNAP Benefits

```{r, echo=FALSE, warning=FALSE}
# This version revised by Cesar
data_long <- weights_geo_inc %>% filter(income_cat == "Less than $35,000") %>% 
  pivot_longer(cols= c(`Monthly mean COVID benefit`, `Monthly mean SNAP benefit`), names_to = "Benefit_Type", values_to = "Dollars_Received")
#View(data_long)
mean_benefit_df <- data_long  %>% group_by(Benefit_Type) %>% dplyr::summarize(Mean=mean(Dollars_Received, na.rm = TRUE))
gg <- ggplot(data = data_long  ) +  
 geom_density(aes(x=Dollars_Received, color=Benefit_Type, fill=Benefit_Type, 
                  text = paste0(
                                'Benefit type: ', Benefit_Type)
                  
                  ), adjust=1.5, alpha=.4
              ) + 
  scale_fill_viridis( option = "D", discrete=TRUE, end=.6) +
  geom_vline(xintercept= unlist(mean_benefit_df[1,2]) , color="darkmagenta", linetype="dashed", linewidth=1) + 
  geom_vline(xintercept= unlist(mean_benefit_df[2,2]), color="darkgreen", linetype="dashed", linewidth=1) + 
  theme_bw() + 
  labs(title="Monthly SNAP Benefit for Households with Income less than $35,000", x="Monthly Average Benefit (Dollars)") 
plotly::ggplotly(gg, tooltip = c("x", "y", "text")) %>%  layout(legend = list(orientation = "h", 
                                     xanchor = "center",
                                     x = 0.5,
                                     y = -0.2) ) 
```

## Box plot

## SNAP Eligibility vs SNAP Loss at Household and Person Levels (Less than $35,000)

```{r, echo=FALSE}
dat = weights_geo_inc %>% filter(income_cat == "Less than $35,000") %>% mutate(people_represented = round(households_represented*average_hh_size, 2))

g <- ggplot(dat, aes(x=SNAP_eligibility_proportion, y=`Monthly mean COVID benefit` - `Monthly mean SNAP benefit`, label = NAME, label2 = households_represented, label3 = average_hh_size, fill = income_cat)) + geom_point() + labs(title = "Monthly Loss of SNAP Benefits per Household per Tract", x = "Proportion of Households Eligible for SNAP", y = "Loss of Monthly SNAP Benefit (Dollars)") + scale_fill_manual(values="#232D4B")
gp <- ggplot(dat, aes(x=SNAP_eligibility_proportion, y=(`Monthly mean COVID benefit` - `Monthly mean SNAP benefit`)/average_hh_size, label = NAME, label2 = people_represented, fill = income_cat)) + geom_point() + labs(title = "Monthly Loss of SNAP Benefits per Person per Tract", x = "Proportion of Households Eligible for SNAP", y = "Loss of Monthly SNAP Benefit (Dollars) / Average Household Size") + scale_fill_manual(values="#232D4B")

ggplotly(g)
```

When looking at potential at-risk tracts in this income category, we examine the tracts with the top 5 highest monthly benefit losses per household and per person. We filter our data to exclude tracts with 5 or fewer households represented in order to consider changes that could impact as many people as possible.

```{r, echo=FALSE}
at_risk_hh = dat %>% filter(households_represented > 5) %>% mutate(`Monthly Benefit Loss` = `Monthly mean COVID benefit` - `Monthly mean SNAP benefit`) %>% arrange(desc(`Monthly Benefit Loss`)) %>% rename(`Average Household Size` = average_hh_size) %>% rename(`Households Represented` = households_represented) %>% select(NAME, `Monthly Benefit Loss`, `Households Represented`, `Average Household Size`) %>% slice(1:5) %>% gt() %>% tab_header(
    title = md("Tracts with Greatest Monthly Benefit Losses per Household"),
    subtitle = md("at income level **Less than $35,000**")
  )
at_risk_hh
```

```{r, echo=FALSE}
hp <- c("Census Tract 4316.01, Fairfax County, Virginia", "Census Tract 4802.05, Fairfax County, Virginia", "Census Tract 4223.02, Fairfax County, Virginia")
highlight_points <- dat %>% filter(NAME %in% hp)

gp <- gp + geom_point(data=highlight_points, aes(x=SNAP_eligibility_proportion, y=(`Monthly mean COVID benefit` - `Monthly mean SNAP benefit`)/average_hh_size), color ="#F84C1E", size=2)
ggplotly(gp)
```

```{r, echo=FALSE}
at_risk_p = dat %>% filter(households_represented > 5) %>% mutate(`Monthly Benefit Loss` = (`Monthly mean COVID benefit` - `Monthly mean SNAP benefit`)/average_hh_size) %>% arrange(desc(`Monthly Benefit Loss`)) %>% rename(`Household Eligibility Proportion` = SNAP_eligibility_proportion) %>% rename(`People Represented` = people_represented) %>% select(NAME, `Household Eligibility Proportion`, `Monthly Benefit Loss`, `People Represented`) %>% slice(1:5) %>% gt() %>% tab_header(
    title = md("Tracts with Greatest Monthly Benefit Losses per Person"),
    subtitle = md("at income level **Less than $35,000**")
  ) %>% 
  tab_style(
    style = list(
      cell_fill(color = "#F84C1E")
    ),
    locations = cells_body(
      rows = c(1, 2, 4))
  )
at_risk_p
```
