---
title: "Overall Policy Effects"
---

```{r, warning=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)
library(acs)
library(tidycensus)
library(tigris)
library(plotly)
library(shiny)
library(ggplot2)
library(viridis)
library(gridExtra)
library(gt)

weights_geo <- read.csv("img/weights_geo.csv")
weights_geo_inc <- read.csv("img/weights_geo_inc.csv")

geo_census <- get_acs(geography = "tract",
                                  year = 2020,
                                  variables = c(Total = "B22003_001"
                                  ),
                                  state = 51, # VA
                                  county = 059,
                                  survey = "acs5",
                                  geometry = TRUE)

geo_census$GEOID <- as.numeric(geo_census$GEOID)
fairfax <- geo_census[, c(1, 2, 6)]

weights_geo <- weights_geo %>% left_join(fairfax, by = c("geoid"="GEOID"))
weights_geo_inc <- na.omit(weights_geo_inc) %>% left_join(fairfax, by = c("geoid"="GEOID"))
```

## New Policy Benefit Loss Effects per Tract

```{r, echo=FALSE, warning=FALSE}

library(sf)
library(leaflet)
library(dplyr)
library(readr)

weights_geo <- weights_geo %>% mutate(Average_benefit_loss = monthly_mean_COVID_benefit - monthly_mean_SNAP_benefit)

#format st
weights_geo <- st_as_sf(weights_geo, crs = 4326) # %>% filter(tot_hh > 10)

#palette colors
pal <- colorBin("viridis", weights_geo$Average_benefit_loss)

#map leaflet

leaflet(weights_geo , options = leafletOptions(attributionControl = FALSE)) %>%
  setView(lng = -77.2464825271027, lat = 38.84619925123701, zoom = 10) %>%  
  addTiles() %>%
  addScaleBar("bottomleft") %>%
  addControl("Average monthly benefit loss", "topright") %>%
  addLegend("bottomright", pal, weights_geo$Average_benefit_loss, opacity = .7) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = pal(weights_geo$Average_benefit_loss), fillOpacity = .7, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"),
    label = paste0(
      "Census tract: ", weights_geo$NAME, "<br/>",
      "Average monthly benefit loss: $", round(weights_geo$Average_benefit_loss,2), "<br/>",
      "Average monthly benefit of SNAP: $", round(weights_geo$monthly_mean_SNAP_benefit,2), "<br/>",
      "Average monthly benefit of SNAP-Covid: $", round(weights_geo$monthly_mean_COVID_benefit,2), "<br/>",
      "Mean income per month: $", round(weights_geo$monthly_mean_income,2) ,  "<br/>",
      "Average household size: ", round(weights_geo$average_hh_size,2) ,  "<br/>",
      "SNAP eligibility (%): ", round(weights_geo$SNAP_eligibility_proportion, 3)
    )     %>%
      lapply(htmltools::HTML)
  ) 

```

## Distributions of COVID and SNAP Benefits at Income Levels

```{r, echo=FALSE, warning=FALSE}
cat_snap <- c('Less than $35,000', '$35,000 to $49,999', '$50,000 to $59,999')
g1 <- ggplot(data=weights_geo_inc %>% filter(income_cat %in% cat_snap) , aes(x=monthly_mean_SNAP_benefit, group=income_cat, fill=income_cat)) +
    geom_density(adjust=1.5, alpha=.5)  + geom_vline(aes(xintercept=mean(monthly_mean_SNAP_benefit)), color="#232D4B", linetype="dashed", linewidth=1) +  geom_density(adjust=1.5, alpha=.5) + scale_fill_viridis(discrete=TRUE, option = "D", end=.9) + theme_bw() + labs(title="Monthly SNAP Benefit for Household Income Levels", x="Monthly Average SNAP Benefit / Household (Dollars)")

g2 <- ggplot(data=weights_geo_inc %>% filter(income_cat %in% cat_snap) , aes(x=monthly_mean_COVID_benefit, group=income_cat, fill=income_cat)) +
    geom_density(adjust=1.5, alpha=.5)  + geom_vline(aes(xintercept=mean(monthly_mean_COVID_benefit)), color="#232D4B", linetype="dashed", linewidth=1) +  geom_density(adjust=1.5, alpha=.5) + scale_fill_viridis(discrete=TRUE, option = "D", end=.9) + theme_bw() + labs(title="Monthly COVID SNAP Benefit for Household Income Levels", x="Monthly Average COVID Benefit / Household (Dollars)")

ggplotly(g2) %>% layout(legend = list(orientation = "h", 
                                     xanchor = "center",
                                     x = 0.5,
                                     y = -0.2) ) 
```

We see the distributions of households at the three income levels shown above, with a mean monthly average COVID SNAP benefit of ~$353 for households making fewer than 60,000 dollars a year.

<br>

```{r, echo=FALSE, warning=FALSE}
ggplotly(g1) %>% layout(legend = list(orientation = "h", 
                                     xanchor = "center",
                                     x = 0.5,
                                     y = -0.2) ) 

```

We see the distributions of households at the three income levels shown above, with a mean monthly average SNAP benefit of ~$237 for households making fewer than 60,000 dollars a year after the policy change.

<br>

## Distributions of COVID-SNAP Changes at Income Levels

```{r, echo=FALSE, warning=FALSE}
g <- ggplot(data=weights_geo_inc %>% filter(income_cat %in% cat_snap) , aes(x=monthly_mean_COVID_benefit-monthly_mean_SNAP_benefit, group=income_cat, fill=income_cat)) +
    geom_density(adjust=1.5, alpha=.5)  + geom_vline(aes(xintercept=mean(monthly_mean_COVID_benefit-monthly_mean_SNAP_benefit)), color="#232D4B", linetype="dashed", linewidth=1) +  geom_density(adjust=1.5, alpha=.5) + scale_fill_viridis(discrete=TRUE, option = "D", end=.9) + theme_bw() + labs(title="Monthly Benefit Loss for Household Income Levels", x="Monthly Average Benefit Loss / Household (Dollars)")

ggplotly(g) %>% layout(legend = list(orientation = "h", 
                                     xanchor = "center",
                                     x = 0.5,
                                     y = -0.2) ) 
```

```{r, echo=FALSE, warning=FALSE}
g <- ggplot(data=weights_geo_inc %>% filter(income_cat %in% cat_snap) , aes(x=(monthly_mean_COVID_benefit-monthly_mean_SNAP_benefit)/average_hh_size, group=income_cat, fill=income_cat)) +
    geom_density(adjust=1.5, alpha=.5)  + geom_vline(aes(xintercept=mean((monthly_mean_COVID_benefit-monthly_mean_SNAP_benefit)/average_hh_size)), color="#232D4B", linetype="dashed", linewidth=1) +  geom_density(adjust=1.5, alpha=.5) + scale_fill_viridis(discrete=TRUE, option = "D", end=.9) + theme_bw() + labs(title="Monthly Benefit Loss for Person at Household Income Levels", x="Monthly Average Benefit Loss / Person (Dollars)")

ggplotly(g) %>% layout(legend = list(orientation = "h", 
                                     xanchor = "center",
                                     x = 0.5,
                                     y = -0.2) ) 
```

## Box plot

```{r echo=FALSE, warning=FALSE}
box3 <- ggplot(weights_geo_inc %>% filter(income_cat %in% cat_snap), aes(x=factor(income_cat, level=c('Less than $35,000', '$35,000 to $49,999', '$50,000 to $59,999')), y=monthly_mean_COVID_benefit - monthly_mean_SNAP_benefit, fill=income_cat)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + labs(title = "Monthly Benefit Loss for Household Income Levels", x = "Income Levels", y = "Estimated Benefit Loss (Dollars)") + scale_fill_viridis(discrete=TRUE, option = "D", end=.9)
ggplotly(box3) %>% layout(legend = list(orientation = "h", 
                                     xanchor = "center",
                                     x = 0.5,
                                     y = -0.2) ) 
```

## SNAP Eligibility vs SNAP Loss at Household and Person Levels (Less than $60,000)

When we consider only the income categories that are impacted by the new policy change that removes COVID SNAP guidelines, we graph the overall impact on each Fairfax census tract for households that make less than $60,000 a year. Each of these points represents a census tract, and each statistic shown at a point is the mean of the statistics for each of the three lowest income levels, weighted against the number of households they represent. 

```{r, echo=FALSE}
names(weights_geo_inc)[names(weights_geo_inc)=='monthly_mean_COVID_benefit'] <- 'Monthly mean COVID benefit'
names(weights_geo_inc)[names(weights_geo_inc)=='monthly_mean_SNAP_benefit'] <- 'Monthly mean SNAP benefit'

dat = weights_geo_inc  %>% filter(income_cat %in% cat_snap) %>% mutate(people_represented = round(households_represented*average_hh_size, 2)) 
```

```{r, echo=FALSE}
under_60 <- dat %>% group_by(geoid) %>% mutate(`Average Household Income` = weighted.mean(monthly_mean_income, households_represented)) %>% mutate(`Monthly Benefit Loss` = weighted.mean(`Monthly mean COVID benefit`, households_represented) - weighted.mean(`Monthly mean SNAP benefit`, households_represented)) %>% mutate(`Proportion Income Lost per Household` = 1 - (weighted.mean(monthly_mean_income, households_represented) - `Monthly Benefit Loss`)/weighted.mean(monthly_mean_income, households_represented)) %>% mutate(`Average Household Size` = weighted.mean(average_hh_size, households_represented)) %>% mutate(`SNAP Eligibility Proportion` = weighted.mean(SNAP_eligibility_proportion, households_represented)) %>% mutate(`Households Represented` = sum(households_represented)) %>% mutate(`People Represented` = sum(people_represented)) %>% distinct(geoid, `Average Household Income`, `Monthly Benefit Loss`, `Proportion Income Lost per Household`, `Average Household Size`, `SNAP Eligibility Proportion`, `Households Represented`, `People Represented`, NAME)

# need to round entire df to 3 digits
#View(under_60)
under_60[,2:6] <- round(under_60[,2:6], 3)
```

```{r, echo=FALSE}
g <- ggplot(under_60, aes(x=`SNAP Eligibility Proportion`, y=`Monthly Benefit Loss`, label = NAME, label2 = `Households Represented`, label3 = `Average Household Size`, label4 = `Proportion Income Lost per Household`)) + geom_point(color="#232D4B") + labs(title = "Monthly Loss of SNAP Benefits per Household per Tract", x = "Proportion of Households Eligible for SNAP", y = "Loss of Monthly SNAP Benefit (Dollars)")

gp <- ggplot(under_60, aes(x=`SNAP Eligibility Proportion`, y=(`Monthly Benefit Loss` / `Average Household Size`), label = NAME, label2 = `People Represented`, label3 = `Proportion Income Lost per Household`)) + geom_point(color="#232D4B") + labs(title = "Monthly Loss of SNAP Benefits per Person per Tract", x = "Proportion of Households Eligible for SNAP", y = "Loss of Monthly SNAP Benefit (Dollars) / Average Household Size")

ggplotly(g)
```

Based on the above graph, we need to examine the tracts that lose the most monthly SNAP benefit as a result of the new policy, as well as which tracts are the most eligible for SNAP in the first place. *The COVID policy affects the range of benefit given but does not greatly impact the eligibility requirements for SNAP, which is why the eligibility proportion does not change as a result of the new policy.*

```{r, echo=FALSE}
at_risk_hh = under_60 %>% ungroup() %>% rename(`Household Eligibility Proportion` = `SNAP Eligibility Proportion`) %>% filter(`Households Represented` > 5) %>% arrange(desc(`Monthly Benefit Loss`)) %>% select(NAME, `Household Eligibility Proportion`, `Monthly Benefit Loss`, `Households Represented`, `Average Household Size`, `Proportion Income Lost per Household`) %>% slice(1:5) %>% gt() %>% tab_header(
    title = md("Tracts with Greatest Monthly Benefit Losses per Household"),
    subtitle = md("For Income under $60,000 a year")
  )
at_risk_hh
```

```{r, echo=FALSE}
at_risk_hh = under_60 %>% ungroup() %>% rename(`Household Eligibility Proportion` = `SNAP Eligibility Proportion`) %>% filter(`Households Represented` > 5) %>% arrange(desc(`Household Eligibility Proportion`)) %>% select(NAME, `Household Eligibility Proportion`, `Monthly Benefit Loss`, `Households Represented`, `Average Household Size`, `Proportion Income Lost per Household`) %>% slice(1:5) %>% gt() %>% tab_header(
    title = md("Tracts with Highest Proportions of SNAP Eligible Households"),
    subtitle = md("For Income under $60,000 a year")
  )
at_risk_hh
```

```{r, echo=FALSE}
ggplotly(gp)
```

We also need to examine the impact of the new policy that reduces benefit on the tracts after considering differing average household sizes per tract. As a result, we look at the most impacted tract by monthly SNAP benefit lost and by eligibility proportion, but this time on a per-person instead of a per-household basis.

```{r, echo=FALSE}
at_risk_p = under_60 %>% ungroup() %>% mutate(`Monthly Benefit Loss` = round(`Monthly Benefit Loss`/`Average Household Size`, 3)) %>% mutate(`Proportion Income Lost per Person` = `Proportion Income Lost per Household`) %>% rename(`Household Eligibility Proportion` = `SNAP Eligibility Proportion`) %>% filter(`Households Represented` > 5) %>% arrange(desc(`Monthly Benefit Loss`)) %>% select(NAME, `Household Eligibility Proportion`, `Monthly Benefit Loss`, `People Represented`, `Proportion Income Lost per Person`) %>% slice(1:5) %>% gt() %>% tab_header(
    title = md("Tracts with Greatest Monthly Benefit Losses per Person"),
    subtitle = md("For Income under $60,000 a year")
  )
at_risk_p
```

```{r, echo=FALSE}
at_risk_p = under_60 %>% ungroup() %>% mutate(`Monthly Benefit Loss` = round(`Monthly Benefit Loss`/`Average Household Size`, 3)) %>% mutate(`Proportion Income Lost per Person` = `Proportion Income Lost per Household`) %>% rename(`Household Eligibility Proportion` = `SNAP Eligibility Proportion`) %>% filter(`Households Represented` > 5) %>% arrange(desc(`Household Eligibility Proportion`)) %>% select(NAME, `Household Eligibility Proportion`, `Monthly Benefit Loss`, `People Represented`, `Proportion Income Lost per Person`) %>% slice(1:5) %>% gt() %>% tab_header(
    title = md("Tracts with Highest Proportions of SNAP Eligible Households"),
    subtitle = md("For Income under $60,000 a year")
  )
at_risk_p
```
